---
title: "White-Britt-ADA-homework-extra"
author: "BA White"
date: "8/30/2020"
output: html_document
---

Add necessary packages

```{r}
library(tidyverse)
library(lme4)
library(AICcmodavg)
```
Challenge 1
• Read the datafiles “gini-index-data.csv” and “world-bank.csv” into two data frames, **gini* and
*worldbank**.
• For the gini data frame, pivot it from wide to long format so that you have a single row with a GINI
index values for each combination of Country Name and Year. HINT: Use {tidyr}’s pivot_longer()
function!
• Print the first 10 lines of each dataset.
```{r}
#Read in datafiles 
gini <- "gini-index-data.csv"
gini <- read_csv(gini)
worldbank <- "world-bank-data.csv"
worldbank <- read_csv(worldbank)

head(gini)
#Now we will pivot the gini data
ginilonger <- gini %>%
  pivot_longer(cols = 5:64)
#Now Print
print(ginilonger)
print(worldbank)
```
CHALLENGE 2
• Join the GINI index and World Bank data sets together using the “Country Code” in gini and “iso3c”
code in worldbank into a new data frame, d. HINT: Use {dplyr}’s left_join() function to add
data from worldbank to every row in gini.
• Convert the variable “Year” to be an integer.
• Convert the variables “Country Name” and “income” to be factors. For “income”, order the levels of
the factor from “Low income” to “High income”. HINT: See the factor() function.
• Drop all rows from d that are missing GINI index data (many years for many countries). HINT: Use
{dplyr}’s filter() function.
• How many rows are in the dataset at this point? HINT: It should be over 1000 and under 2000!
```{r}
#Join datasets
newdf <- left_join(ginilonger, worldbank, by = c("Country Code" = "iso3c"))
glimpse(newdf)

#Convert "Year" ie "name" to integer
newdf$name = as.integer(as.character(newdf$name))
newdf$'Country Name' = as.factor(as.character(newdf$'Country Name'))
glimpse(newdf)


###????
#order levels of factor from low to high income
#unique(newdf$income)
#newdf$income  <- factor(newdf$income, levels = c("Low income", "Lower middle income", "Upper middle income", "High income"))
#glimpse(newdf)
###

#drop missing GINI Index missing values
d <- newdf %>%
 filter(!is.na(value))

#How many rows? 
glimpse(d)
```
CHALLENGE 3
• Using {ggplot}, with the data frame, d, as your data argument, plot on the same graph the GINI index
across years for the “Group of Seven” (G7) countries, which are the largest IMF-advanced economies in
the world (i.e., the United States, United Kingdom, Canada, France, Germany, Italy, and Japan) plus
China, Russia, and Iran. HINT: When you assign the data= argument in {ggplot}, you can have the
assignment be the result of a piped set of {dplyr} operations.
• NOTE that in d at this point, the names “Russia” and “China” are not used as country names (check
out the country codes RUS and IRN). Use {dplyr} and str_replace() to change the original country
names to “Russia” and “Iran” before making your plot.
```{r}
##change the names of Russia and Iran(?) 
d$country <- str_replace(d$country,"Iran, Islamic Rep.", "Iran")
d$country <- str_replace(d$country,"Russian Federation", "Russia")

##use ggplot to plot the GINI Index/years for largest IMF advanced economies in the world, ie. the "Group of Seven Countries" 
d %>% 
filter(country %in% c("United States", "United Kingdom", "Canada", "France", "Germany", "Italy", "Japan", "China", "Russia", "Iran")) %>%
  ggplot(aes(name, value, color = country)) +
  geom_point() +
  geom_smooth(se = FALSE, method = lm)

```
CHALLENGE 4
• Again using {ggplot}, with d as your dataset, add Brazil, Bolivia, Colombia, India, Vietnam, Rwanda,
Uganda, and Ethiopia to the list of countries you examine and make a plot with four panels, where you
facet by “income”.
NOTE: For both Challenge 4 and Challenge 5, try to generate the plots with a SINGLE line of code,
and orient the legend at the bottom and remove the title of the legend. HINT: You will probably
want to use theme(legend.position="bottom", legend.title = element_blank()).
```{r}
#using ggplot, add Brazil, Bolivia, Colombia, India, Vietnam, Rwanda, Uganda, and Ethiopia
d %>% 
filter(country %in% c("Brazil", "Bolivia", "Colombia", "India", "Vietnam", "Rwanda", "Uganda", "Ethiopia", "United States", "United Kingdom", "Canada", "France", "Germany", "Italy", "Japan", "China", "Russia", "Iran")) %>%
  ggplot(aes(name, value, color = country)) +
  geom_point() +
  geom_smooth(se = FALSE, method = lm)+
  theme(legend.position = "bottom", legend.title = element_blank())

```

CHALLENGE 5
• Using {dplyr}’s group_by() and summarize() functions, create a new data frame, gini_summary,
that includes the minimum, maximum, average, oldest GINI index, latest GINI index, year of the oldest
GINI index, and year of the latest GINI index for all countries in d, as well as a variable, “gini_change”,
that is the difference in the GINI index between the most recent and the oldest GINI indices for a
country divided by the number of years between the most recent and oldest indices (i.e., that summarizes
the change in GINI index across time for each country, given the data available). Arrange that table in
descending order of the “gini_change” within each “income” group, with the “income” groups ordered
from highest to lowest.
HINT: This can all be done in single line of code by piping d into group_by(), summarize(),
and arrange() functions. Within summarize(), you will be creating multiple new summary
variables. Here’s one way that you can create one of these variables, the GINI value for the oldest
GINI index in the dataset for a particular group
oldest_gini = `GINI Index`[which(`Year` == min(`Year`))]
3
```{r}
```
CHALLENGE 6
• Using {ggplot} and the gini_summary dataset, make a boxplot of “gini_change” by “income” group.
```{r}
```

CHALLENGE 7:
• Return to data frame d and filter() it to only include the 18 countries mentioned in Challenges 3
and 4.
• Using linear mixed effect models, explore how the GINI index is related to “population”, “income”, and
“gdp_capita” as predictors when “Country Name” is included as a random factor in the model. You
should construct a full model that includes all three predictors plus the random factor of “Country
Name” as well as a complete set of nested models (no need to include interaction terms). The null
model should include only the random factor. Once you have constructed a full set of models, construct
an table ranking the models based on AICc. Which is the best supported model? What other models
are included in the set contributing to a cumulative AIC weight of >0.90? Which predictors appear in
all of the top models?
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
```
